\chapter{SIMD Programming}

With SIMD defined, the question which naturally follows is how dedicated SIMD
instruction sets can be put into use. Unfortunately, there is no simple answer.
The particular machine architecture you choose to target affects what
possibilities are available. 


\section{Writing SIMD Code}

\subsection{Compiler Intrinsics}
Compiler intrinsics are essentially built-in functions which compile down to
specific CPU instructions, or short sequences thereof. These are primarily
available for the C programming language, which transitively makes the available
for using within C++. Other fully compiled languages, such as Rust, and even
some partially compiled languages such as C\# have facilities which exist to
fill the same niche. For both x86 and ARM, there exist compiler intrinsics for
every SIMD instruction, allowing for a level of granularity in choosing
instructions comparable to using assembly, while still retaining the benefits
from using a compiler, such as register allocation and peephole optimization.

\subsection{(Inline) Assembly}
Another possible approach is through the use of assembly. As would be expected,
this approach offers the more control over the result. 

\subsection{SIMD Libraries}
A third possible method of leveraging a CPU's SIMD capabilities would be to make
use of libraries which abstract over SIMD instruction sets. These libraries
generally rely on compiler intrinsics and/or inline assembly to construct thin
abstraction layers which provide a uniform interface over a .

\subsection{Auto-vectorization}
Arguably the easiest, although least effective method of taking using SIMD
instructions is to rely on auto-vectorization. This refers to a family of
compiler optimizations which recognize opportunities.

\subsection{SPMD Compilers}
There is a particular category of compilers which are available.

\subsection{x86 Intrinsics}
When using compiler intrinsics to program x86 CPUs, the most valuable resource
will be
\href{https://www.intel.com/content/www/us/en/docs/intrinsics-guide/index.html}{Intel
Intrinsics Guide}.

x86's compiler intrinsics are originally from Intel's ICC and ICP compilers, now
the ICX and ICPX compilers. This does not mean that you need to use Intel's
compiler to make use of these intrinsics. These intrinsics are also offered by
GCC, Clang, and MSVC. Therefore, compiler intrinsics are relatively portable
across compilers.



\subsection{ARM Intrinsics}
